---
title: "Assign2TEMPLATE"
output:
  pdf_document: default
---

### 1. Z ~ N(0,1)
```{r}
# A. P(Z ≤ 0.79)
pnorm(0.79)

# B. P(Z ≤ -0.23)
pnorm()

# C. P(Z > 1.42)
# D. P(-0.23 ≤ Z ≤ 1.10)
# E. Find the value z such that P(Z ≤ z) = 0.2843
# F. Find the value z such that P(Z > z) = 0.3264


```
### 2. Y ~ N(5.2, 0.8)
```{r}
#A P(Y < 6)

#B P(Y > 5) 

#C P(4.8 < Y < 6.8)

#D y such that P(Y < y) = 0.85                                    

```
### 3. Property of Y ~ skewed distn (mu=80, sd=5)
A.  Give an interval with the property that at least 75% of the data will be in that interval.
What rule did you use to determine the interval?

B.  Describe the distribution of . Give the mean, standard deviation and shape of the
distribution. (3 pts)


### 4. Seeds: A random sample of n=25 seeds from a particular bean population is obtained. The weight of each seed is recorded. 
```{r, fig.height= 3, fig.width= 4, fig.align='cnter'}
Seeds<-read.csv("/Users/bensharp/Dropbox/STAT511Spring2020/Assignments/Assign2/Seeds.csv")
str(Seeds)
# A. Construct a histogram of the data. Also give the sample mean and sample standard deviation. (3 pts)
 

# B. Give a 95% confidence interval for μ (population mean seed weight).

```
C. Interpret your confidence interval from part B.

D. Do you think the CI is valid? In other words, are assumptions satisfied?
 

#5 Describe how the following affect the width of the confidence interval (assuming everything else is held constant). Answer should be increase, decrease or stays the same.
A. Sample size increases.
B. Confidence level increases.
C. Standard deviation increases.


