---
title: "Assign 5"
author: "Natalie Schmer"
header-includes: \usepackage{bm}
output: word_document
editor_options:
  chunk_output_type: console
---

## 1.  
Refer to Problem 6.42 which deals with lung capacity of rats exposed to ozone. Note: For consistency, please calculate the differences as After – Before for all questions. (12 pts)

```{r, message=FALSE}
library(tidyverse)

#load data and arrange data to Rats and Rats2
Rats <- read.csv("/Users/natalieschmer/Desktop/GitHub/stats_511/data/ASCII-comma/CH06/ex6-42.TXT", quote ="'")

str(Rats)
```

### 1A.  
Are the differences normally distributed? Support your answer by including a qqplot of differences in your assignment.
```{r, out.height="250px"}

Rats <- Rats %>%
            mutate(differences = After - Before)

qqnorm(Rats$differences)
qqline(Rats$differences)
```

**Based on the qqplot and the fact that the points generaly fall along the 1:1 line, the differences do appear normally distributed**


### 1B.  
Is there sufficient evidence to support that ozone exposure increases lung capacity? Use the paired t-test with $\alpha$=0.05. State the hypotheses, test statistic, p-value and conclusion. (4 pts)
```{r}
t.test(Rats$differences)
```

**Hypotheses: Null hypothesis is that the mean difference in lung capacities is 0, and the alternative is that the mean difference in lung capacities is NOT 0. t = 3.885, p-value = 0.002541. Since p < 0.05, we reject the null hypothesis and conclude the true mean in difference of lung capacity is NOT 0.**

### 1C.  
Estimate the size of the increase in lung capacity after exposure and construct a 95% t-based CI. Note: provide a standard “two-sided” CI here.
```{r}
t.test(Rats$differences)
```
**The size of the increase in lung capacity after exposure is about 1.21 and the 95% CI is (0.5237735, 1.8928932) **


### 1D.  
Rerun the test from part B using the Wilcoxon Paired (Signed Rank) test. Give your p-value and conclusion. Use the wilcoxsign_test() function from the coin package with distribution = “exact”. Remember to use a one-sided alternative. (4 pts)
```{r}
# below code just opens the help for these two functions
?wilcox.test

library(coin)
?wilcoxsign_test

wilcoxsign_test(After ~ Before, data = Rats, distribution = "exact", alternative = "greater")
```
**In this case, p-value = 0.002441, so much les than 0.05, and we can still conclude that the true mean in difference of lung capacity is NOT 0, and is greater than 0.**

## 2  
Designs for comparing means using College Board data on SAT scores.  State whether the following is testing: 1) a single mean versus a claim in null hypothesis,  2) Means from two independent samples, or 3) Means from matched pairs.  Also state the degrees of freedom associated with the t-test. (1 pt for hypotheses of test, 1 pt for df => 10 total))

### 2A.  
**Hypothesis is testing a single mean versus a claim in null hypothesis, and df = 49**

### 2B.
**Hypothesis is testing means from two independant samples, and df = 50**

### 2C.
**Hypothesis is testing means from two independant samples, df= 45.**

### 2D.
**Hypothesis is testing means from matched pairs, df= 24.**

### 2E.
**Hypothesis is testing means from matched pairs, df = 19**

## 3.  
Refer to problem 7.6  Vehicle Speeds Single Standard Deviation   

### 3A.  
Construct a 95% CI for $\sigma$.  
```{r}
s=11.35
n=100
```


```{r}
df <- 99
s <- 11.35

upper <- sqrt((df*s^2)/(qchisq(.025, 99)))
lower <- sqrt((df*s^2)/(qchisq(.975, 99)))
```

### 3B.  
Using $\alpha$=0.05, test $H_A$: $\sigma$ > 10. Give your test statistic, rejection rule and conclusion. (4pts)  
```{r}
x2 <- (99*(s^2))/ 10^2

pval_3b <- 1-pchisq(x2, df= 99)
```
**The test statistic is 127.53, and p = 0.0282775. Since p < 0.05, we reject the null hypothesis and conclude that the true standard deviation is > 10.**

### 3C.  
**The distributional assumptions include being a random sample, independent observations, normally distributed data.**

### 3D.  
**Even though 10 is in our confidence interval, we still reject the null because our p value is less than 0.05 **

