---
title: "Assign 11"
author: "Natalie Schmer"
output:
  word_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

## 1. Bank Salaries
```{r setup1}
library(ggplot2)
library(knitr)
library(tidyverse)
saldata <- read.csv("/Users/natalieschmer/Desktop/GitHub/stats_511/data/BankSalaries.csv")
Large <- subset(saldata, Size=="Large")  
Small <- subset(saldata, Size=="Small")
```

### 1A. Scatter Plot
```{r out.height="180px"}
ggplot(data = saldata, aes(x = LOS, y= Wages, shape = Size))+
  geom_point(size = 2)+
  geom_smooth(mapping = aes(LOS, Wages, color = Size), method = "lm", se = T)+
  theme_bw()
```

### 1B Regression Models and Equation for line
```{r}
lm_large <- lm(Wages ~ LOS, data = Large)
#y = 49.45 + 0.05595(LOS)

lm_small <- lm(Wages ~ LOS, data = Small)
#y = 35.87192 + 0.10416(LOS)
```

### 1C. Interpretation  
For large banks, wages start higher but for one unit increase in LOS increases by 0.056, whereas for small banks wages start lower but for one unit increase in LOS increase by 0.10.

### 1D.Test for slope
```{r}
summary(lm_large)

summary(lm_small)
```
*For large banks, p = 0.282 and for small banks, p = 0.0001712. For, large banks, there is not evidence that LOS is linearly related to wages, but there is evidence this is true for small banks. *

### 1E CI for Intercept
```{r}
confint(lm_large, level = 0.95)
confint(lm_small, level = 0.95)
```
*Ultimately, an employee would be better off at a large bank because the representing the starting wage, starts higher than a smaller bank.*

### 1F. CI for Mean LOS=96
```{r}
newdata <- data.frame(LOS = 96)
predict(lm_large, newdata, interval = "predict", level = 0.95)
predict(lm_small, newdata, interval = "predict", level = 0.95)

predict(lm_large, newdata, interval = "confidence", level = 0.95)
predict(lm_small, newdata, interval = "confidence", level = 0.95)
```
*After 8 years, an emplyee would still be better off at a larger bank, as the range of possible wages as represented by the confidence interval is higher than for smaller banks. *


### 1G. Outlier
```{r}
car::outlierTest(lm_large)
#row 15 of the large dataset, LOS is 70 and wage is 97.68
```

### 1H. T-test
```{r, out.height="150px"}
boxplot(LOS~Size, data =saldata)

t.test(LOS~Size, data =saldata)
```
*Since the boxplots have a fair amount of overlap and p > 0.05, we cannot conclude that LOS is significantly different between bank sizes.*



## Q2 Steel Quadratic 
```{r setup2}
Steel<-read.csv("/Users/natalieschmer/Desktop/GitHub/stats_511/data/Steel.csv")
str(Steel)
```

### 2A. (**4pts**)  
```{r,echo=FALSE, out.height="150px"}
lm_2a <- lm(Strength ~ Thick, Steel)
par(mfrow = c(1, 2))
plot(lm_2a, which = c(2, 1))
```
*Based on the plots, the regression assumptions do not seem to be well met. the Residuals vs fitted show deviations away from the 0 line, and the QQ plot is not very linear.*

### 2B.  (**4pts**)   
```{r}
regfit2a <- anova(lm_2a)

ANOVA2a <- lm(Strength ~ as.factor(Thick), Steel)
ANOVAfit2a <- anova(ANOVA2a)

#lack of fit
anova(lm_2a, ANOVA2a)

#F 
regfit2a$'Sum Sq'

SSreg = regfit2a$'Sum Sq'[2]
SSanova = ANOVAfit2a$`Sum Sq`[2]
DFreg = regfit2a$Df[2]
DFanova = ANOVAfit2a$Df[2]
F = ((SSreg-SSanova)/(DFreg-DFanova))/ANOVAfit2a$`Mean Sq`[2] 

1-pf(F,(DFreg-DFanova),DFanova) # p-value
```
*Since p< 0.05, there is evidence of a lack of fit.*

### 2C. (4 pts)
```{r, fig.width = 4.5, fig.height = 3.5}
ggplot(data = Steel, aes(x = Thick, y = Strength))+
  geom_point()+
  geom_smooth(method = "lm", formula = y ~ x + I(x^2))+
  theme_bw()

summary(lm(Steel$Strength ~ Steel$Thick + I(Steel$Thick^2)))
```

## Q3 SAT Scores
```{r setup3, echo=FALSE}
sat <- read.csv("/Users/natalieschmer/Desktop/GitHub/stats_511/data/SAT.csv")
```

### 3A.
```{r, out.height="200 px"}
sat_cor <- sat %>% 
              select(-year)

pairs(sat_cor)
```

### 3B.
```{r}
cor(sat_cor, method = "pearson")
```
*The strongest correlation is between female verbal and male verbal*

### 3C
```{r}
cor.test(sat_cor$Female.Verbal, sat_cor$Female.Math, method = "pearson")
```
*Since p > 0.05, and the correlation coefficient is very low, there is little to no correlation bewteen female verbal and female math scores.*

