---
title: "Assign9"
author: "Natalie Schmer"
output:
  word_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---
28 points total, 2 points per problem part unless otherwise noted.

## Q1 Toxic Tomatoes

```{r}
#A 
#New
new_1a <- 5/50

#Old
old_1a <- 9/50

#B 
prop.test(c(5, 9), c(50, 50), correct = T)
#X-squared = 0.74751, p-value = 0.3873

#C
#set up matrix
tomatoes <- matrix(c(5, 45, 9, 41), byrow = T, nrow = 2)
colnames(tomatoes) <- c("Toxic", "Non-toxic")
rownames(tomatoes) <- c("New", "Old")
tomatoes

#test
(tomato_test <- chisq.test(tomatoes, correct = T))
#Z-squared = 0.748, p-value = 0.3873

#D
tomato_test$expected

#E 
fisher.test(tomatoes)
#p-value = 0.3881
```

F. **Based on part D, fisher's test is preferred because the cell counts are very small as compared to the overall sample size.**
```{r}
#G
prop.test(c(5, 9), c(50, 50), correct = T, alternative = "less")
```

## Q2 Anesthesia
```{r}
#A
#DrugA
a_2 <- 22/47

#DrugB
b_2 <- 13/47 

#B (4 pts)
#Since the sample sizes are small, fischer test 
drugs <- matrix(c(4, 18, 9, 16), byrow=T, nrow =2)
rownames(drugs) <- c("Drug A Yes", "Drug A No")
colnames(drugs) <- c("Drug B Yes", "Drug B No")
drugs

fisher.test(drugs) 
#odds ratio = 0.4029197, p-value= 0.2071
```

## Q3 Case Control Study
3A.
```{r}
bird <- (98/199)/(1-(98/199))
nobird <- (141/469)/(1-(141/469))
odds_ratio <- bird/nobird
```
**The bird group has higher odds of getting lung cancer by about 2.25:1**

3B. 
```{r}
birds <- matrix(c(328, 141, 101, 98), nrow = 2, byrow = T)
colnames(birds) <- c("Healthy control", "Cancer patients")
rownames(birds) <- c("No Bird", "Bird")
birds

epitools::oddsratio(birds, method = "wald")
```
**Based on the interval, it does appear that there is a relationship between bird ownership and lung cancer, where bird owners are more likely to have lung cancer. These estimates are significantly different than 1, since 1 is not in the CI.**

3C. 
```{r}
(bird_test <- chisq.test(birds, correct = T))
```
**Since the p value is much less than 0.05, we reject the null hypothesis that the proportion of lung cancer patients is the same between bird owners and those that didn't own birds, and conclude that there is a greater rate of lung cancer with bird owners.**
