---
title: "Assign2TEMPLATE"
author: "Natalie Schmer"
output:
  word_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

### 1. Z ~ N(0,1)
```{r}
# A. P(Z ≤ 0.79)
pnorm(0.79)

# B. P(Z ≤ -0.23) 
pnorm(-0.23)

# C. P(Z > 1.42)
1- pnorm(1.42)

# D. P(-0.23 ≤ Z ≤ 1.10)
# same as = P(Z<= 1.1) - P(Z<= -0.23)
pnorm(1.1) - pnorm(-0.23)

# E. Find the value z such that P(Z ≤ z) = 0.2843
qnorm(0.2843)

# F. Find the value z such that P(Z > z) = 0.3264
qnorm(1-0.3264)

```
### 2. Y ~ N(5.2, 0.8)
```{r}
{
  mu <- 5.2
  sigma <- 0.8
}

#A P(Y < 6)
pnorm((6-mu)/sigma)

#B P(Y > 5) 
1-pnorm(((5-mu)/sigma))

#C P(4.8 < Y < 6.8)
pnorm((6.8-mu)/sigma) - pnorm((4.8-mu)/sigma)

#D y such that P(Y < y) = 0.85 
mu + (qnorm(0.85)*sigma)

```
### 3. Property of Y ~ skewed distn (mu=80, sd=5). Random sample of n = 100 is drawn. 

A.  Give an interval with the property that at least 75% of the data will be in that interval.

**80 +/- 10, so (70, 90)**

What rule did you use to determine the interval?
**I used the part of Chebyshev's rule that says at least 75% of the data will lie within Y(bar) +/- 2s**

B.  Describe the distribution of Y(bar) Give the mean, standard deviation and shape of the
distribution. (3 pts)
**As stated, the mean is 80, the standard deviation is 5, and the shape of the distribution is normal due to the Central Limit Theorem.** 

### 4. Seeds: A random sample of n=25 seeds from a particular bean population is obtained. The weight of each seed is recorded. 
```{r, fig.height= 3, fig.width= 4, fig.align='cnter'}
seeds<-read.csv("~/Desktop/Seeds.csv")
str(seeds)
# A. Construct a histogram of the data. Also give the sample mean and sample standard deviation. (3 pts)
hist(seeds$Weight, 
     col="grey", 
     main="Histogram of Seeds", 
     xlab="Seed Weight")

mean(seeds$Weight)

sd(seeds$Weight)

# B. Give a 95% confidence interval for μ (population mean seed weight).
t.test(seeds$Weight, 
       mu= mean(seeds$Weight), 
       conf.level = .95)

```
C. Interpret your confidence interval from part B.
**There is a 95% chance that the mean value will fall between [479.1754, 573.0646]**

D. Do you think the CI is valid? In other words, are assumptions satisfied?
**I think the CI is valiid. Even though the histogram looks slightly skewed to the left, the distribution given by the t-test is still generally right even if the data are not completely normal.**

#5 Describe how the following affect the width of the confidence interval (assuming everything else is held constant). Answer should be increase, decrease or stays the same.
A. Sample size increases. 
**CI width decreases**
B. Confidence level increases.
**CI width increases**
C. Standard deviation increases.
**CI width increases**


