---
title: "Assign1Template"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

This is a template for completing the first assignment.  It is not required that you use it, but hopefully helps with organziation and to see some of the functionality of R Markdown.  You'll need to tailor this file for your own use.  For example, the grader need not have this paragraph included.  

#1.
Use the data described in Problem 3.30 regarding lumber. From the files you downloaded above, you will find the data under CH03, named ex3-30.txt. Use the following commands to import and summarize the data.

```{r, setup}
#Use path name with R Markdown
lumber <- read.csv("~/Documents/GitHub/CSU/stats_511/data/assignment_1/ASCII-comma/CH03/ex3-30.TXT", quote = "'")

#View(lumber)  this doesn't work well with R Markdown
str(lumber)
```

**A. Include the histogram in your assignment.**
```{r}

hist(lumber$Number, 
     main = "Histogram of Lumber",
     xlab = "Lumber",
     ylab = "Frequency")


```
**B. Give the mean and median of the sample.**
```{r}

(mean(lumber$Number))
# 7.728571

median(lumber$Number)
# 8 
```
**C. Does the data appear to be “normal” (bell-shaped)? Justify your response based on your histogram from above.**  

Yes, I think the data could be considred normal due to the fact that there is only 1  majority of the observations fall in the 7-9 range, which matches the median and mean, with decreasing frequency moving away from the middle. 
<br>

#This is another style or example of submitting a question via R Markdown (a bit more compact) 

#2 Survival Times

```{r}
survival <- read.csv("~/Dropbox/STAT511Fall2019/Assignments/OL_Data/CH03/ex3-7.TXT", quote = "'")
str(survival)

survival <- read.csv(file.choose(), quote = "'")
View(survival)
str(survival)
mean(survival$StandardTherapy)
sd(survival$StandardTherapy)
boxplot(survival)

A. What is the sample mean and sample standard deviation for each of the therapies?
B. Construct side-by-side boxplots showing the survival times for each therapy.

#A
#Standard 

#New

#B

```




