---
title: "Assign 4"
author: "Natalie Schmer"
output:
  word_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---
```{r, include=FALSE}
library(boot)
library(tidyverse)
library(BSDA)
```


## 1.
### 1A
**The null hypothesis is that the mean process time is 15 minutes, and the alternative hypothesis is that the mean process time is less than 15 minutes**

### 1B
```{r}
{
  n <- 30
  test.stat <- -1.81
  side2_p <- 0.08 
}

pnorm(test.stat, lower.tail = T)
```

## 2.  
### 2A 
Find power for n = 20.
```{r}
{
  sd <-  2
  true_mean <- 16 #hyp= 15
  n <- 20 
}

power.t.test(n = n, 
             delta = 1, 
             sd = sd, 
             sig.level = 0.05, 
             type = "one.sample", 
             alternative = "one.sided")
```
**The power is 0.6951493**

### 2B Larger Conjecture for std dev 
```{r}
power.t.test(n = n, 
             delta = 1, 
             sd = 3, 
             sig.level = 0.05, 
             type = "one.sample", 
             alternative = "one.sided")
```
**Lower power with higher sd**


### 2C Larger Sample.
```{r}
power.t.test(n = 30, 
             delta = 1, 
             sd = sd, 
             sig.level = 0.05, 
             type = "one.sample", 
             alternative = "one.sided")
```
**Higher power with larger sample size**

### 2D Smaller Alpha.
```{r}
power.t.test(n = n, 
             delta = 1, 
             sd = sd, 
             sig.level = 0.01, 
             type = "one.sample", 
             alternative = "one.sided")
```
**Lower power with higher alpha**

### 2E Larger Delta Conjecture. 
```{r}
#If conjectured mean was 17, delta would be 2 
power.t.test(n = n, 
             delta = 2, 
             sd = sd, 
             sig.level = 0.05, 
             type = "one.sample", 
             alternative = "one.sided")
```
**Higher power with larger delta**

### 2F Find n for power = .90.
```{r, message=FALSE}

power.t.test(power = 0.9, 
             sd= sd,
             delta = 1,
             alternative = "one.sided",
             type = "one.sample")
```
**n = 35.65268 which rounds to 36**

## 3. 
### Use the data from Problem 5.29 which deals with lead concentrations in estuarine creeks.
```{r}
PB <- read.csv("/Users/natalieschmer/Desktop/GitHub/stats_511/data/ASCII-comma/CH05/ex5-29.TXT", quote ="'")
```

### 3A  
```{r, fig.height=3.5}

hist(PB$Lead, main= "Lead", xlab= "Lead concentration in estuarine creeks")
qqnorm(PB$Lead)
qqline(PB$Lead)

(shapiro.test(PB$Lead))
```
**The SW test of normality gives a p-value much less than 0.5, so the data is not normal. This is backed up by the histogram, which is heavilty skewed to the left, and the qq plot/ qq line which is not close to straight**


### 3B
```{r}
summary(PB$Lead)
```
**Mean is 37.24, median is 32.00**

### 3C
```{r, message=FALSE}
BSDA::SIGN.test(PB$Lead, md = 30)$p.value
```
**P-value is greater than 0.05, so we fail to reject the null hypothesis**

### 3D 
```{r}
BSDA::SIGN.test(PB$Lead, md = 30)
```
**Upper Achieved CI (17.0000 , 41)**

### 3E
```{r}
t.test(PB$Lead)
```
**Standard 95% CI (24.86550, 49.62099)**

### 3F
```{r}
#Define the function
mean.fun <- function (d, i)
{ m <- mean(d[i])
  n <- length(i)
  v <- (n-1)*var(d[i])/n^2
  c(m, v)
}

set.seed(3700)
results2 <- boot::boot(data = PB$Lead, mean.fun, R = 1000)
boot::boot.ci(results2, type = "all")
```
**Studentized 95% CI: (28.03, 56.33)**

###3G  
**For a cumulative assessment, the mean would work the best**

## 4.
```{r}
Jets <- read.csv("/Users/natalieschmer/Desktop/GitHub/stats_511/data/ASCII-comma/CH06/ex6-43.TXT", quote = " ' ")
```

### 4A
```{r, fig.height=3, fig.width=4}
#Boxplot for when data are in 2 columns
boxplot(Jets)
```  

### 4B
```{r}
mean(Jets$NarrowBodied, na.rm = TRUE)
sd(Jets$NarrowBodied, na.rm = TRUE)

mean(Jets$WideBodied)
sd(Jets$WideBodied)

```

### 4C
```{r}
#reshape data 
jets_long <- Jets %>%
                pivot_longer(cols= c("WideBodied",   "NarrowBodied"),
                            names_to = "jets",
                             values_to = "body",
                             values_drop_na = F)

(jets_t <- t.test(body ~ jets, 
       data = jets_long,
       var.equal = T))

(jets_t$conf.int) 
```
**Based on the CI, we can conclude there is a difference on populations because 0 is not in the CI**

### 4D
```{r}
(sd(Jets$NarrowBodied, na.rm = TRUE)/sd(Jets$WideBodied))
```

**Based on part B, (recalculated in 4D), we can use a pooled variance t-test because smax/smin < 2**

### 4E
```{r}
(jets_t_4e <- t.test(body ~ jets, 
       data = jets_long,
       var.equal = T, 
       mu = 0,
       alternative = "two.sided"))

jets_t_4e$p.value
```
**Since the p-value is < 0.05, we can reject the null hypothesis that the noise level is the same bewteen the two groups**
